---
title: "Installation"
date: "Last modified: `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This sections describes how to install the `{mapme.agriculture}` package. 
Because the package handles some spatial data, we assume that you have
a working installation of [GDAL](https://gdal.org/index.html) and [PROJ](https://proj.org/) 
on your system. Please refer to the respective installation guides to learn about 
how to install these programs on your OS.

In order to be able to download data from the FAO [WaPOR API](https://wapor.apps.fao.org/home/WAPOR_2/1) make sure to register
an account and create and API key that you should have at hand to identify
your queries to the server.
 
Installing the `{mapme.agriculture}` package is then achieved trough the following
function call: 


```{r installation, eval = F}
remotes::install_github("mapme-initiative/mapme.agriculture")
```

The installation process will install potential missing R dependencies.
After the successful installation, you are ready to use the package's 
functionality. Simply load the package for your R session with the library
call:

```{r setup, eval = F}
library(mapme.agriculture)
```

