---
title: "Calculate Productivity Indicators"
date: "Last modified: `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculate Productivity Indicators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Getting Started

```{r setup, warning=F, message=F}
library(mapme.agriculture)
library(sf)
library(terra)
library(dplyr)
```

In this documentation we are presenting how to use the mapme.wapor package
to download and process data sets from the [FAO WaPOR](https://wapor.apps.fao.org/home) 
API to calculate several indices to assess biomass water productivity.
WaPOR makes available several layers at different resolutions concerned with
agricultural productivity. While this package can be used do download
several of these layers, we will focus on the ones that are required to 
calculate supported productivity indices. 
These product are:

- Total Biomass Productivity (TBP): This data layer indicates the total biomass
production (kg/ha) for two distinct seasons every year
- Phenology (PHE): This data layer indicates the decade the start and end of season
in a given pixel was detected. It is used to aggregate other indicators according
to the length of the respective season
- Land Cover Classifcation (LCC): This data layer delivers a broad classification
scheme of land cover /land use and it is used to focus the analysis on specific
land covers
- Transpiration (T): This data layer indicats decadal transpiration values (mm)
for each pixel. It is used to calculate the total amount of transpired water by
plants during the length of the growing season
- Gross Biomass Water Productivity (GBWP): This data layer indicates the amount
of water used in the production of biomass (kg/m³). It is a gross estimate because
besides plant transpiration also soil evaporation and interception is included.
The package will calculate net biomass water productivity based on the transpiration
layer


## Download input data 

Above mentioned products are needed for the other functions to 
work properly. We can use the download function to download other products as well, 
however, this is out of scope for this vignette. In the code chunk below we show 
how the download utility can be used to query a spatio-temporal domain for 
download. Note that you will need a personal API key that can be acquired in the 
user section of the WaPOR [website](https://wapor.apps.fao.org/home/WAPOR_2/1).


```{r warning=F, echo=T, message=F, eval = F}
aoi = st_read(system.file("extdata", "aoi.gpkg", package = "mapme.agriculture"))
aoi = st_transform(aoi, 4326) # transform to WGS84 because the API expects lat/lon coordinates
collection = "WAPOR_2" # wapor colletion 2.0
products = list("TBP" = "L2_TBP_S",
                "PHE" = "L2_PHE_S",
                "LCC" = "L2_LCC_A", 
                "T" = "L2_T_D",
                "GBWP" = "L2_GBWP_S") # The products we want to download

begin = as.Date("2009-01-01") # begin date is inclusive
end = as.Date("2015-01-01") # end date is exclusive
APIkey = "<-your key goes here->" # can be obtained in the profile section of the WAPOR website 
dir.creat("data")

for(i in 1:length(products)){ # download products in a loop
  wapor_queryRaster(collection = collection,
                    product = products[[i]],
                    aoi = aoi,
                    begin = begin,
                    end = end, 
                    APIkey = APIkey, 
                    outdir = system.file("extdata", "input", package = "mapme.agriculture"), 
                    cutline = FALSE, 
                    tiled = TRUE, 
                    compressed = TRUE, 
                    overviews = TRUE,
                    sleep_time = 2) # increase sleep time for larger extents to avoid being banned 
}

```

The above function call will download the queried products to the data directory.
Note that downloading larger raster extents can take a while because the FAO server
will need to process the query. Make sure to increase the sleep time of the function
to an appropriate value to avoid querying the server to frequently for the current status.


## Calculation of NBWP

With the package we deliver some previously downloaded raster files to showcase
how we can process these. The very first step would be to calculate the Net Biomass
Water Productivity, that is the ratio between biomass produced to transpiration only.
For this we provide the wapor_nbwp function. Note that NBWP can only be calculated
when data for transpiration from the previous year as well as the complete following
year is available. The reason is that, depending on where we are on the globe, a 
growing season might start in the previous year and it might as well extend to the 
following year. That is also why the earliest year we can calculate NBWP is 2010, 
because WaPOR transpiration data is available from 2009 onwards. The function
will cause an error either if a year before 2010 is queried or if the last year
is later than the previous year at the time the function is called. Besides the
numeric vector indicating the years we are interested in we need to supply the
function with a single character vector of the input files. Note that the names 
of these files follow a certain standard so they should not be changed after download.
Additionally, you should specify an output directory where two new directories are
created. In the first step the function sums up the seasonal total sum of transpiration
and outputs these rasters to a directory called "T_S". In the next step the seasonal
total biomass production is divided by the seasonal transpiration to get NBWP. These
rasters are written to a directory called "NBWP".

```{r calc-nbwp}
input_files = list.files(system.file("extdata", "input", package = "mapme.agriculture"), recursive = T, pattern = ".tif", full.names = T)
dir.create(system.file("extdata", "intermediate", package = "mapme.agriculture"), showWarnings = F)
nbwp_files = wapor_nbwp(input_files = input_files,
                        years = 2010:2014, 
                        outdir = system.file("extdata", "intermediate", package = "mapme.agriculture"),
                        verbose = T)


```

We can take a look at the NBWP files for the first season.

```{r plot-nbwp}
plot(rast(nbwp_files[grep("S1", nbwp_files)]))
```


## Calculate productivity indicators

In the next step we are interested in calculating a number of indices from the
NBWP as well as the input files. The supported indices are the following:

- land cover statistics: mode and median of landcover, year of breakpoint, frequency of 
specified land cover classes
- the presence of a double season in a certain pixel
- yearly values of the raw values of TBP, GBWP, NBWP
- Global Land Productivity: Based on TBP, GBWP, and NBWP, the ratio a pixel
achieves in a certain year in comparison of the 95%-quantile across all years
- Yearly Land Productivity: the same as above except that the threshold value
is calculated on the basis of the 95%-qunatile for each year
- differences between epochs: the raw parameters and the productivity indices
are averaged for two epochs and the difference between these two values is returned
- linear trends for the raw parameters and the productivity indicators: each
raster has two layers, the first is the estimate of the the trend, the second the p-value

We can use the wapor_indices function to calculate the complete raster set that is written
to the specifed output directory. The land cover classification map can be used to
restrict the calculation to certain land cover types. When mask_values is omitted
the raster files are not masked prior to calculation. Note that we can specify
the number of cores for parallel computation. However, it is only advised to use
this functionality for very large rasters. For example, when calculating 
the indices based on the rasters delivered with this package any parallel processing
will substantially increase computation time because the extent is rather small.

```{r calc-indices}
input_files2 = c(input_files, nbwp_files)
index_files = wapor_indices(input_files2, 
                            years = 2010:2014, 
                            mask_values = c(20,30), 
                            ncores = 1, 
                            epoch1 = 2010:2012, 
                            epoch2 = 2013:2014, 
                            outdir = system.file("extdata", "output", package = "mapme.agriculture"),
                            overwrite = TRUE)
```

Let's visualize both of the land productivity indictators for one of the selected
land cover classes. We will start with visualizing the global indicator.

```{r plot-global}
plot(rast(index_files[grep("20-nbwp-s1-lp1_global.tif", index_files)]))
```

Now we will move to visualize the same parameter but this time it is based on a
yearly threshold.

```{r plot-yearly}
plot(rast(index_files[grep("20-nbwp-s1-lp1_yearly.tif", index_files)]))
```

Differences are subtle, but note that these indicators were calculated only for
pixels where the respective land cover classification was found to correspond to 
a value of 20. We can quickly cross-check this by visualizing the input
land cover classification.

```{r plot-lcc}
lcc = rast(input_files[grep("LCC", input_files)])
lcc[lcc !=  20] = NA
plot(lcc)
```



## Extract zonal statistics

Since we have now the calculated indices in the next step we are interested in
extracting this information and summarizing it per polygon. To showcase this 
functionality we included an additional vector data set with a number of polygons
in the study area.

```{r plot-aoi}
aoi2 = st_read(system.file("extdata", "aoi2.gpkg", package = "mapme.agriculture"))
plot(lcc[[1]]) 
plot(aoi2, add = T)
```


For the extraction, we simply specify the path to the index files as well as the
aoi to the wapor_extract function.

```{r extract-indices}
extracted_values = wapor_extract(index_files, aoi2)
```

This will give us a list with different tibble objects representing different
levels of the indicator extraction. From the output below we see that
in the first object we find the main land cover statistcs. In the second object,
the land cover frequency for the specified mask values is found. In the third
object we get the averaged yearly values of the raw parameters as well as 
the productivity indicators. In the fourth object the number of double seasons
is indicated while the difference in double seasons between epochs is found in the
fifth object. For the raw parameters and the productivity indicators the difference
between epochs are found in the sixth object while the trend parameters are found in the
last one.

```{r structure}
str(extracted_values, 2)
```


Note, that all returned tibbles are formatted in a long format. The column ID
specifies the index of in relation to the inputted aoi object. Thus if we wish
to combine a certain table with the original aoi object we can proceed as follows:

```{r join}
aoi2$ID = 1:nrow(aoi2)
left_join(aoi2, extracted_values$indices)
```

