---
title: "Workflow"
date: "Last modified: `r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.height = 5
)
```

In the following, the most important steps of the indicator calculation with
of `{mapme.agriculture}` are described; the figure below guides the reader through these
descriptions. The goal of this tutorial is to use the data provided by FAO's WaPOR
portal to calculate a number of indicators concerned with agricultural productivity. 
Because a key data set, Net Biomass Water Productivity (NBWP), is not provided
at a spatial resolution of 100 meters, its calculation is an important step
for any analysis. Subsequently, a number of different indicators area automatically
calculated for either all pixels in the study area of for user-defined land-use 
classes. The output are a number of raster files, each representing one of the
indicators. Optionally, zonal statistics for field locations can be extracted
which are handed over as multiple sf-objects that can be used in further analysis.

-  query and download basic input data sets from the WaPOR API for a given spatio-temporal extent
-  calculate the sum of seasonal transpiration and use it to calculate NBWP
-  calculate raster sets of a number of selected indicators potentially restriced to certain land-use classes
-  optionally, extract zonal statistics from these indicators to use them in further analysis

```{r, echo = FALSE}
library(DiagrammeR)

mermaid("
  graph TB

    classDef default fill:#c2dddd,stroke-width:2px,stroke:#034858;
 
    A(Query & download input data)
    B(Calculate seasonal NBWP)
    C(Calculate prodictivity indicators)
    D(Extract zonal statistics)
    A-->B
    B-->C
    C-->D
")
```

